---
title: "Data Preparation"
author: "PRU"
date: "02/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(arrow)
library(dplyr)
library(purrr)
library(tidyr)
library(sf)
library(sftrack)
library(geosphere)
library(trrrj)
library(ggplot2)
library(patchwork)

# apdf read util functions
source("./R/construct_apt_fn.R")
source("./R/read_csv12.R")
```

## Validation Data Set from APDF

Data preparatory step: 

* reading in APDF data downloaded from PRISME
* trimming data set to data points needed for this study
* writing out March-May for 2019, 2020, and 2021 for EGLL, LFPG, and EHAM

Following write out, chunk/extracting deactivated with eval=FALSE.

```{r apdfval, eval=FALSE}
#' read apdf data extracted from PRISME, subset for required columns
#'
read_apdf <- function(.apt, .yr){
  # my filename
  fn <- construct_apt_fn(.apt, .yr)
  # read only needed columns
  colspec <- cols_only(
     AP_C_FLTID = col_character()
    ,AP_C_REG   = col_character()
    ,ADEP_ICAO  = col_character(), ADES_ICAO = col_character()
    ,MVT_TIME_UTC = col_character()   # could not get col_datetime() going
    ,ARCTYP     = col_character()  
    ,AC_CLASS   = col_character()
    ,AP_C_RWY   = col_character()
    ,C40_CROSS_TIME = col_character()
    ,C40_CROSS_LAT = col_double() , C40_CROSS_LON = col_double()
    ,C40_CROSS_FL = col_integer(), C40_BEARING = col_double()
    ,C100_CROSS_TIME = col_character()
    ,C100_CROSS_LAT = col_double() , C100_CROSS_LON = col_double()
    ,C100_CROSS_FL = col_integer(), C100_BEARING = col_double()
    ,SRC_PHASE  = col_character()
  )
  # read and make nicer names
  df <- read_csv12(fn, colspec) %>% 
    rename( FLTID = AP_C_FLTID, REG = AP_C_REG
           , ADEP = ADEP_ICAO, ADES = ADES_ICAO
           , MVT_TIME = MVT_TIME_UTC, TYPE = ARCTYP
           , WTC = AC_CLASS, RWY = AP_C_RWY
           , C40_BRG = C40_BEARING, C100_BRG = C100_BEARING) %>%
    mutate(across(contains("TIME"), lubridate::dmy_hms))  # coerce timestamps
  return(df)
}

#' from given data set extract study months, i.e. March (,April), and May .
extract_study_month <- function(.df, .months = c(3:5)){
  df <- .df %>%
    mutate(MOF = case_when(
      !is.na(MVT_TIME) ~ lubridate::month(MVT_TIME)
      ,is.na(MVT_TIME) & !is.na(C40_CROSS_TIME) ~ lubridate::month(C40_CROSS_TIME)
      ,is.na(MVT_TIME) & is.na(C40_CROSS_TIME) & !is.na(C100_CROSS_TIME) ~ lubridate::month(C100_CROSS_TIME)
      ,TRUE ~ as.double(NA)
    )) %>% 
    filter(MOF %in% .months) %>%
    select(-MOF)
}

extract_and_save_apdf_valdata <- function(.apt, .yr){
  df <- read_apdf(.apt, .yr) %>%
    extract_study_month()
  write_csv(df, file = paste0("./data/val-apdf-", .apt, "-", .yr,".csv.gz"))
}

# testing (apdf <- read_apdf("EGLL", 2019) %>% extract_study_month())
apts <- c("EGLL", "EHAM","LFPG")
yrs  <- c(2019:2021)

expand.grid(apt = apts, year = yrs) %>%
  purrr::pwalk(.f = ~ extract_and_save_apdf_valdata(..1, ..2))
```

read in validation data set

```{r}
fns <- list.files(path = "./data/", pattern = "^val-apdf-", full.names = TRUE)

colspec <- cols(
  FLTID = col_character(),
  REG = col_character(),
  ADEP = col_character(),
  ADES = col_character(),
  MVT_TIME = col_datetime(format = ""),
  TYPE = col_character(),
  WTC = col_character(),
  RWY = col_character(),
  SRC_PHASE = col_character(),
  C40_CROSS_TIME = col_datetime(format = ""),
  C40_CROSS_LAT = col_double(),
  C40_CROSS_LON = col_double(),
  C40_CROSS_FL = col_double(),
  C100_CROSS_TIME = col_datetime(format = ""),
  C100_CROSS_LAT = col_double(),
  C100_CROSS_LON = col_double(),
  C100_CROSS_FL = col_double(),
  C40_BRG = col_double(),
  C100_BRG = col_double()
)

# read in and tag (with source airport) validation data
read_apdf_val <- function(.fn, .colspec){
    my_apt <- stringr::str_extract(.fn, pattern = "[A-Z]{4}")
    val_ds <- read_csv(.fn, col_types = .colspec) %>%
        mutate(APT = my_apt) %>%
        select(APT, everything())
    return(val_ds)
}

apdf <- purrr::map_dfr(.x = fns, .f = ~ read_apdf_val(.x, colspec))
```

```{r}
tmp <- apdf %>% 
  filter(SRC_PHASE == "ARR") %>%
  mutate( DOF  = lubridate::date(MVT_TIME)
         ,YEAR = lubridate::year(MVT_TIME)
         ) %>%
  group_by(APT, YEAR, DOF) %>%
  summarise(ARRS = n(), .groups = "drop") %>%
  na.omit() 

tmp$DOF2 <- tmp$DOF
lubridate::year(tmp$DOF2) <- 666
  
p <- tmp %>%
  ggplot() +
  geom_line(aes(x = DOF2, y = ARRS, colour = APT )) +
  facet_grid(APT ~ YEAR) +
  theme_minimal() +
  labs(y = NULL, x = NULL) +
  theme(legend.position = "none")
# ggsave("./figures/airport-timeline-2019-2020-2021.png", dpi = 320, height = 6, width = 8)
p
```




```{r}
ds <- read_feather("./data-raw/EGLL_2019-03-01_history.feather") %>% as_tibble()
glimpse(ds)
```
```{r}
names(ds)
```

TODO - via icao24 we can get aircraft type

```{r}
make_standard_names <- function(.df){
  df <- .df %>%
    rename( TIME   = timestamp
           ,ICAO24 = icao24
           ,FLTID  = callsign
           ,ADEP   = origin
           ,ADES   = destination
           ,LAT    = latitude
           ,LON    = longitude
           )
  return(df)
}
ds <- ds %>% make_standard_names()
```


create some meta data

```{r}
meta <- ds %>% 
  group_by(FLTID, ICAO24) %>% 
  summarise(N = n(), ADEP = unique(ADEP), ADES = unique(ADES)
            ,ENTRY_TIME = min(TIME), END_TIME = max(TIME)
            ,ELAPSED_TIME = difftime(END_TIME, ENTRY_TIME, units = "mins")
            ,.groups = "drop")

meta
```


```{r}
meta %>% 
  ggplot() + 
  geom_point(aes(x = N, y = as.numeric(ELAPSED_TIME))) +
  labs(title = "data points versus elapsed time")
```

>TODO: function to extract trajectory payload (or use {sftrack/traj})
> force ordering: group_by(callsign, icao24) %>%
    arrange(time, .by_group = T)

```{r}
set.seed(666)
arrs <- meta %>%
  mutate(ID = paste0(FLTID, ICAO24)) %>%
  filter(ADES == "EGLL") %>%
  slice_sample(n = 10)

ds %>% filter(FLTID %in% arrs$FLTID, ICAO24 %in% arrs$ICAO24) %>%
ggplot(aes(x = LON, y = LAT, colour = ICAO24)) +
  geom_point()
```

```{r}
ds %>% filter(FLTID %in% arrs$FLTID, ICAO24 %in% arrs$ICAO24) %>%
ggplot(aes(x = TIME, y = altitude, colour = ICAO24)) +
  geom_point()
```

> TODO - smooth altitude     

```{r}
ds %>% filter(FLTID %in% arrs$FLTID, ICAO24 %in% arrs$ICAO24) %>%
  select(FLTID, ICAO24, TIME, LAT, LON, altitude) %>%
  rename(longitude = LON, latitude = LAT) %>%
  smooth_positions() %>%

ggplot(aes(x = TIME, y = altitude, colour = ICAO24)) +
  geom_point() +
  labs(title = 'smoothing adds more outliers for altitude as it pulls good data!' )
```

```{r}
ds %>% filter(FLTID %in% arrs$FLTID, ICAO24 %in% arrs$ICAO24) %>%
  select(FLTID, ICAO24, TIME, LAT, LON, altitude) %>%
  group_by(FLTID, ICAO24) %>% 
  mutate( v_change = altitude - lead(altitude)
         ,jump     = abs(v_change) > 500
        # , jump2 = lead(jump), jump3 = lag(jump)
         , alt_sm  = zoo::rollmedian(altitude, k = 7, align = "center", fill = NA)) %>%

ggplot(aes(x = TIME, y = alt_sm, colour = ICAO24)) +
  geom_point() 
```






OBSERVATION - trajectories feel to be cut off at runway or close to ground := no ground data

```{r}
ds %>% filter(ICAO24 == "3991e4") %>%
  ggplot(aes(x = LON, y = LAT, colour = FLTID)) +
  geom_point()
```
```{r}
tmp <- ds %>% 
  filter(ICAO24 == "3991e4", FLTID == "AFR23GC") %>% 
  filter(TIME > max(TIME) - 600) 

p1 <- tmp %>% ggplot(aes(x = LON, y = LAT)) + geom_point()
p2 <- tmp %>% ggplot(aes(x = TIME, y = geoaltitude)) + geom_point()

p1 + p2
```

runway system 
lookup rather than calc: https://www.pgc.umn.edu/apps/convert/

Runway 09L/27R
Dimensions:	12802 x 164 feet / 3902 x 50 meters
Surface:	Hard
	Runway 09L	Runway 27R
Coordinates:	N51°28.65' / W0°29.10'	N51°28.66' / W0°26.00'
Elevation: 	79 	78
Runway Heading: 	091° 	271°
Displaced Threshold: 	1004 Feet 

Runway 09R/27L
Dimensions:	12008 x 164 feet / 3660 x 50 meters
Surface:	Unknown
	Runway 09R	Runway 27L
Coordinates:	N51°27.89' / W0°28.94'	N51°27.90' / W0°26.04'
Elevation: 	75 	77
Runway Heading: 	091° 	271°
Displaced Threshold: 	1007 Feet

```{r}
aip<- tribble(
   ~ICAO , ~REF, ~NAME,          ~TYPE, ~ID,  ~LAT,      ~LON,      ~GEOM_ALT
  ,"EGLL", "ARP", "Aerodr. Ref", "ARP", "ARP",51.4706,   -0.461941, 83
  ,"EGLL", "09L", "RWY 09L/27R", "THR", "R1", 51.4775,   -0.485,    79
  ,"EGLL", "27R", "RWY 09L/27R", "THR", "R1", 51.477667, -0.433333, 78
  ,"EGLL", "09R", "RWY 09R/27L", "THR", "R2", 51.464833, -0.482333, 75
  ,"EGLL", "27L", "RWY 09R/27L", "THR", "R2", 51.465,     -0.434,   77
)
```


TODO ===========================

LFPG
Latitude 	49.0127983093
Longitude 	2.54999995232
Geom. Altitude 	392ft (119m)

EHAM
Latitude 	52.3086013794
Longitude 	4.763889789579999
Geom. Altitude 	-11ft (-3m)

```{r}
tmp <- ds %>% filter(ADES == "EGLL") %>% 
  group_by(ICAO24, FLTID) %>% 
  filter(TIME >= max(TIME) - 30)
  
 aip %>% filter(TYPE == "THR") %>%  
    group_by(ID) %>% 
    mutate(  LAT2 = lead(LAT), LON2 = lead(LON) ) %>% 
   na.omit() %>%
   
  ggplot() + 
   geom_segment(aes(x = LON, xend = LON2, y = LAT, yend = LAT2)
                , colour = "green", size = 5) + 
   #---------- ARP
   geom_point( data = aip %>% filter(REF == "ARP")
              , aes(x = LON, y = LAT), colour = "red", size = 3) + 
   #---------- last positions
   geom_point(data = tmp, aes(x = LON, y = LAT)) +
   #---------- trim
   xlim(-0.65, -0.35) + ylim(51.455, 51.485)
```

Note: interesting offset of some arrivals !?!

## define airport box

bboxfinder: 51.461066,-0.506401,51.482666,-0.412846

```{r}
egll_box <- st_bbox(c( xmin = -0.506401, xmax = -0.412846
                      ,ymin = 51.461066, ymax = 51.482666)
                    ,crs = st_crs(4326)) %>%
  st_as_sfc()
```

pick a test trajectory

```{r}
trj <- ds %>%
  filter(ICAO24 == "a14bdd", FLTID == "AMX007") %>%
  filter(TIME >= max(TIME) - 200)
```

Utility functions to coerce lat/lon to sf points or linestring

```{r}
cast_latlon_to_pts <- function(.df, .crs = 4326, .drop_coord = TRUE){
  pts_sf <- .df %>%
    sf::st_as_sf(coords = c("LON","LAT"), crs = .crs, remove = .drop_coord)
  return(pts_sf)
}

cast_pts_to_ls <- function(.pts_sf, .group_var){
  ls_sf <- .pts_sf %>%
    group_by({{ .group_var}}) %>%
    summarise(do_union = FALSE, .groups = "drop") %>%
    st_cast("LINESTRING")
  return(ls_sf)
}

cast_latlon_to_ls <- function(.df, .crs = 4326, .drop_coord = TRUE, ...){
  pts_sf <- cast_latlon_to_pts(.df, .crs, .drop_coord)
  ls_sf  <- cast_pts_to_ls(pts_sf, .group_var = NULL)
}
```

create a nested data set of flights with the "latest" 4 mins (~ 240 secs)

```{r}
rq <- ds %>% 
  select(TIME, ICAO24, FLTID, ADEP, ADES, LAT, LON) %>% 
  group_by(ICAO24, FLTID, ADEP, ADES) %>% 
  filter(TIME >= max(TIME - 240)) %>%
  tidyr::nest()
```

determine / confirm arrivals := trajectory (linestring) intersects airport box

```{r}
confirm_flight_phase <- function(.df, .apt_box){
  df <- .df %>%
    mutate(ARR = map(
        .x = data
       ,.f = ~ cast_latlon_to_ls(.x) %>%
                    st_intersects(.apt_box, sparse = FALSE))
    ) %>%
    unnest(ARR)
  df <- df %>% mutate(ARR = as.logical(ARR))
}

rq <- rq %>% confirm_flight_phase(egll_box)
rq
```


## Identification of landing runway

The identification of the landing direction is based on a simple geospatial heuristic.    
Conceptually, aircraft are aligned with the runway (centerline) before landing.
An aircraft is assigned to a landing runway based on the closeness of its pre-landing positions to the extended runway centerline.

```{r}
trj <- ds %>%
  filter(ICAO24 == "a14bdd", FLTID == "AMX007") %>% 
  filter(TIME >= max(TIME) - 300)

trj_pts <- trj %>% cast_latlon_to_pts()

trj %>%
  ggplot() + 
  geom_point(aes(x = LON, y = LAT))
```

extended runway centerlines

```{r}
rwy_ctr_line <- function(.rwy_df = aip, .ctrl_length = 10000){
  df <- .rwy_df %>% 
    filter(REF != "ARP") %>% 
    select(REF, NAME, LAT, LON) %>% 
    group_by(NAME) %>% 
    mutate( LAT2 = case_when(!is.na(lag(LAT))  ~ lag(LAT)
                            ,!is.na(lead(LAT)) ~ lead(LAT))
           ,LON2 = case_when(!is.na(lag(LON))  ~ lag(LON)
                            ,!is.na(lead(LON)) ~ lead(LON))
          ) %>% 
  # calculate "reverse" runway bearing with geosphere::bearingRhumb
  mutate( RBRG  = bearingRhumb(p1 = cbind(LON2, LAT2), p2 = cbind(LON, LAT))
         )
  
  # determine "endpoint" of extended centerline at d = 10000 meters
  tmp <- with(df, destPointRhumb(cbind(LON, LAT), b= RBRG, d = .ctrl_length)) %>% 
    as_tibble() %>% rename(LON3 = lon, LAT3 = lat)
  
  # combine and return
  df <- df %>% bind_cols(tmp)
  return(df)
}

rwy_ctr_line_pts <- function(.rwy_ctr_line, .debug = FALSE){
  # could not get pivot_longer work with multiple cols
  tmp1 <- .rwy_ctr_line %>% select("REF":"LON")
  
  # include opposite runway threshold
  tmp2 <- .rwy_ctr_line %>% select("REF","NAME", "LAT2","LON2") %>%
    rename(LAT = LAT2, LON = LON2)
  
  # centerline end point determined beforehand
  tmp3 <- .rwy_ctr_line %>% select("REF","NAME", "LAT3","LON3") %>%
    rename(LAT = LAT3, LON = LON3)
  
  if(.debug == TRUE){ 
    df <- bind_rows(tmp1, tmp2, tmp3)
  }else{
    df <- bind_rows(tmp1, tmp3)
  }
  
  df <- df %>% arrange(REF, NAME)
  return(df)  
}
```

showcase rwy centerline(s) and sample arrival

```{r}
rwy_ctr_line(aip) %>% rwy_ctr_line_pts() %>% 
  cast_latlon_to_pts() %>% cast_pts_to_ls(REF) %>% 
  
  ggplot() + 
  geom_sf(aes(colour = REF), linetype = "dashed") + 
  geom_sf(data = trj_pts)
```

based on centerline we can associate the - estimated - landing runway

As a heuristic we check for points within a certain distance from the extended centerline.    
To select one of several (potentially crossing, less well spaced) runways the number of hits (close to runway) is determined.

```{r}
# spatial join (aka "overlay" with "within")
# to identify most likely ~ closest runway (extended centerline-string)
estimate_rwy <- function(.trj_pts, .rwy_ctrl_ls, .ctrl_offset = 100){
  # stop if nots
tmp <- st_join(.rwy_ctrl_ls, .trj_pts
               , join = st_is_within_distance, dist = .ctrl_offset) %>%
  st_drop_geometry() %>%        # keep just the "spatially joined" dataframe
  na.omit() %>%                 # eliminate no hits
  group_by(REF) %>%
  summarise(N = n(), .groups = "drop") %>%
  mutate( TOT   = sum(N, na.rm = TRUE)       # na.rm should not be necessary
         ,TRUST = N / TOT)                   # share in case several runways
  
return(tmp)
}
```

```{r}
rwy_ctr_line_ls <- aip %>% 
  rwy_ctr_line() %>% 
  rwy_ctr_line_pts() %>% 
  cast_latlon_to_pts() %>% cast_pts_to_ls(REF)

estimate_rwy(trj_pts, rwy_ctr_line_ls)
```

test for rq data set

```{r}
rq <- rq %>% filter(ADES == "EGLL") %>%
  mutate( rwy = map(.x = data, .f = ~estimate_rwy(.x %>% cast_latlon_to_pts() , rwy_ctr_line_ls))
         ,RWY = map(.x = rwy, .f = ~.x %>% filter(TRUST == max(TRUST)) %>% pull(REF))
  ) %>% tidyr::unnest(RWY)
```

TODO - chase mismatches
with(rq, table(ARR, RWY))
       RWY
ARR     27L 27R
  FALSE   2  11
  TRUE  343 278

## Determine Actual Landing Time

Random checking of landing time (MVT_TIME, SRC_PHASE == "ARR") from APDF with max(TIME) of ds shows rounding to minute in reporting of EGLL. 
It appears that the traffic library provides - given ground coverage at Heathrow - landing times - at least for runway direction east, i.e. 27R/L on 1. March.

> TODO - check delta over whole data set - this saves quite some time


Heuristic: if we need to detect landing and determine landing time

* check coverage on final - for wider out, take characteristic point (e.g. 1000m & 1500m) before threshold, determine groundspeed & apply compression slow down for GPI around 300m inward from threshold
* ....


### visual check of throughput (~ landing time APDF ~ trajectory last point time)

check on 15 min bins shows relatively good match; OSN overlaps APDF

>TODO - check whether difference are non-ADSB flights

```{r}
tmp1 <- apdf %>% 
  filter(APT == "EGLL", SRC_PHASE == "ARR") %>% 
  select(FLTID, REG, ADES, MVT_TIME) %>%
  mutate(BIN = lubridate::floor_date(MVT_TIME, unit = "15 minute")) %>%
  group_by(BIN) %>% 
  summarise(N = n()) %>%
  mutate(SRC = "APDF")

tmp2 <- ds %>%
  filter(ADES == "EGLL") %>%
  select(FLTID, ICAO24, TIME) %>%
  group_by(FLTID, ICAO24) %>%
  filter(TIME == max(TIME)) %>%
  mutate(BIN = lubridate::floor_date(TIME, unit = "15 minute")) %>%
  group_by(BIN) %>% 
  summarise(N = n()) %>%
  mutate(SRC = "OSN")

tmp1 %>%
  filter(BIN %in% tmp2$BIN) %>%
  bind_rows(tmp2) %>%
  ggplot() +
  geom_step(aes(x = BIN, y = N, colour = SRC)) +
#  facet_wrap(.~ SRC) +
  NULL
```

## Characteristics Points - Intersection with Distance from Ref Point (e.g. ARP)

library trrrj to the rescue!

```{r}
arp  <- c( lon = aip %>% filter(REF == "ARP") %>% pull(LON)
          ,lat = aip %>% filter(REF == "ARP") %>% pull(LAT)
          )

c40  <- polygon_at_distance(geo = arp, d = 40)
c100 <- polygon_at_distance(geo = arp, d = 100)
c200 <- polygon_at_distance(geo = arp, d = 200)

trjs    <- ds %>% mutate(ID = paste0(FLTID, ICAO24)) %>% filter(ID %in% arrs$ID)
trjs_ls <- trjs %>% select(TIME, FLTID, ICAO24, ID, LAT, LON) %>% 
  cast_latlon_to_pts() %>% 
  cast_pts_to_ls(.group_var = FLTID)

# to calculate intersections between polygon (circle) and linestring (trjectory)
# polygon needs to be casted to a "linestring" (otherwise we get what is inside the poly)
crossings <- st_intersection(trjs_ls, c40 %>% st_cast("LINESTRING"))

ggplot() + 
  geom_sf(data = c200) +
  geom_sf(data = c100) +
  geom_sf(data = c40) +
  geom_sf(data = trjs_ls, aes(color = FLTID)) +
  geom_sf(data = crossings, color = "red", size = 2)
  
```

I struggled for a while to find a good way to determine the crossing time.    
The following is a brute force technique

```{r}
extract_pos_at_distance_from_point <- function(.trj_df, .pt, .dist, .col_name = CX){
  dist_NM_m <- .dist * 1852
  pos <- .trj_df %>%
    mutate(DIST = geosphere::distHaversine(cbind(LON, LAT), .pt)) %>%
    filter(between(DIST, dist_NM_m - 200, dist_NM_m + 200 )) %>%
    group_by(FLTID) %>%
    mutate(DIST2 = abs(DIST - dist_NM_m))%>%
    filter(DIST2 == min(DIST2)) 
  pos <- pos %>% rename({{.col_name}} := DIST)
  return(pos)
}
```

```{r}
trjs %>% select(FLTID, ICAO24, TIME, LAT, LON) %>% 
  extract_pos_at_distance_from_point(arp, 40, DIST_C40)
```

alternative implementation with {sftrack}

> TODO: could not get the as_sftrack/traj() going for df - took detour via df -> sf -> sftraj

```{r}
#library(sftrack)

cast_latlon_to_sftrjs <- function(.df_trjs) {
  sftrjs <- .df_trjs %>%
    select(FLTID, ICAO24, TIME, LAT, LON) %>%
    cast_latlon_to_pts() %>%
    as_sftraj( coords=c("LON","LAT")
            ,group = c(id = "FLTID", month = "ICAO24")   # {sftrack} expects month
            , time = "TIME") 
  return(sftrjs)
}

sf_trjs <- trjs %>% cast_latlon_to_sftrjs()
head(sf_trjs, 10)
```

```{r}
sf_trjs[st_intersects(sf_trjs, c40 %>% st_cast("LINESTRING"), sparse = FALSE),]
```

## Playing around with grid cells - h3r

H3 is a hexagonal hierarchical geospatial indexing system. 

```{r}
library(h3)

coords <- c(37.3615593, -122.0553238)
resolution <- 8
# Convert a lat/lng point to a hexagon index at resolution 8
(h3_index <- geo_to_h3(coords, resolution)) 
```

assigning coordinates to cells is a breeze!!!!

```{r}
trjs_to_cells <- trjs %>% select(FLTID, ICAO24, TIME, LAT, LON) %>% 
  mutate(CELL = geo_to_h3(c(LAT, LON), res = 8))

trjs_to_cells %>% slice_sample(n = 15)
```

Check how many hits we get for resolution 8 (a bit under 500m radius)

```{r}
trjs_to_cells %>% group_by(FLTID, CELL) %>% summarise(N = n())
```

```{r}
cell_counts <- trjs_to_cells %>% group_by(CELL) %>% summarise(N = n())
hexs <- h3_to_geo_boundary_sf(cell_counts$CELL)
hexs <- hexs %>% bind_cols(cell_counts)
hexs
```

```{r}
ggplot() +
  geom_sf(data = hexs, aes(color = as.factor(N)))
```

```{r}
ggplot() +
  geom_sf(data = hexs, aes(color = as.factor(N), fill = as.factor(N))) +
  coord_sf(xlim = c(-0.75,0.5), ylim = c(51.2,51.62))
```


what do we need per (arrival) trajectory?

- identify flight (entry 200NM to airport)      DONE (by traffic lib)
- unique id (ensure no same FLTID and ICAO24)
- landing runway           DONE
- actual landing time      DONE ~ taken from given flights
- asma points
  - 40NM crossing time, crossing lat/lon & bearing
  - 100NM crossing time, lat/long & bearing
- associated time to fly
- associated distance to fly
- associated grid cell
